---
title: 'Generating data from random variables'
output: word_document
date: "2026-03-09"
---

# Generate data 

In R we have these options:

- d density
- p for probability
- q for quantiles 
- r random generation of data

to generate dta from the normal inivariate distribution we use **rnorm**

```{r}
#rnorm
set.seed(7369)
n <- 1000
mu <- 0
sigma <- 1
z <- rnorm(n, mean = mu, sd = sigma)
```
### Data Description
We summerize the relized data


```{r}
require(skimr)
skim_without_charts(z)
```

We use the transformation
$$
Y_i = \mu + \sigma z, \:\ i=1, \ldots,n
$$
We have that $Y \sim N(\mu, \sigma^2)$
Yi = mu + sigma Z

We generate data related to height of a certain group of people where we except the average is around 140 with a standard deviation of about 6.

```{r}
y <- 140 + 6*z
```
Report the descriptive statistics
```{r}
skim_without_charts(y)
```


Report some figures and describe the data 

```{r}
hist(y, breaks =30,
     main = "Realized values from Y~N(140,36)",
     prob = TRUE,
     xlim = c(110,173),
     ylim = c(0,0.08),
     ylab = "Density",
     col = "yellow1")
```

*Comments expected on the shape and in particular highest frequencies and tails behavior reporting some observed values.* 

## Plot

Report some figures and describe the data: 

### Histigram

### Empirical cumulative distribution function

```{r}
plot(ecdf(y),
     main = "Empirical cumulative distribution Y~N(140,36)")
```

*Fix some values of interest in the x-axis and reports some comments on the related observed probability.*

### Quantile-quantile plot

```{r}
qqnorm(y, col = "blue", main="Y~N(140,36)")
```

