---
title: "Multiple linear regression model"
output: word_document
date: "2026-03-23"
---

```{r}
load("profits.Rdata")
```

y: profit margins
x1: revenues 
x2: share of investments

```{r}
pairs(profits, 
      panel = panel.smooth,
        main = "Scatterplot matrix ")
```
We estimate the model using the lm function 

```{r}
lm1 <- lm(y ~ x1 + x2, data = profits)
lm1 <- lm(y ~ ., data = profits)

```


We make a short description of the results through the summary function

```{r}
summary(lm1)
```

```{r}
require(scatterplot3d)
ex1 <- scatterplot3d(profits[,2],
                     profits[,3],
                     profits[,1],
                     type = "h", 
                     pch = 20)
ex1$plane3d(lm1, col = "red")
```



# Residuals


```{r}
lm1$residuals[1]
```


Observed values $y_i$

```{r}
profits$y[1]
```
Fitted value $\hat{y}_i$

```{r}
lm1$fitted.values[1]
```


$z_i = y_i -  \hat{y}_i$

```{r}
profits$y[1] - lm1$fitted.values[1]
```


## Detection of influential observations

We plot residuals alone


```{r}
plot(lm1$residuals, pch = 16, cex = 0.4)
```


## Residuals against fitted value 

```{r}
plot(lm1$fitted.values, lm1$residuals, pch = 16, cex = 0.4)
```



## Residuals against covariate 

```{r}
plot(profits$x1, lm1$residuals, pch = 16, cex = 0.4)
```


```{r}
library(car)
residualPlots(lm1)
```


# Empirical distribution function of residuals

We plot also the residuals to see their distribution

```{r}
plot(ecdf(lm1$residuals))
```


We compare the shape of the empirical function with a theoretical distribution of reference which is the student t distribution with (n-p-1) degrees of freedom 

```{r}
plot(ecdf(lm1$residuals))
n <- dim(profits)[1]
p <- 2
df <- (n - p -1)
curve(pt(x,df), col = "red", add = T)
```

# Leverage points

We are going to calculate first the values of the H matrix

```{r}
lev <- hatvalues(lm1)
skimr::skim_without_charts(lev)
```

From description of the values there are no problems of leverages since they are ranging near by zero and the variability is low.

The plot is done with the function inside the library named faraway

```{r}
require(faraway)
halfnorm(lev)
```

No leverage values we can however inspect the observations referred to the highest values related to id 148 and 409

```{r}
profits[148,]
profits[409,]
```

# Cook's distance

```{r}
cook <- cooks.distance(lm1)
skimr::skim_without_charts(cook)
```

```{r}
halfnorm(cook, pch = 16, cex = 0.4)
```

If you detect leverages, and/or outliers and/or influential values check the id in the data and try to estimate the model removing once each time.


```{r}
summary(lm1)
```

so $s^2$ is 
```{r}
0.09639^2
```


Realized estimate of the variance supposed for the error terms in the model.

```{r}


```










