---
title: '#Example15'
output: html_document
date: "2026-04-08"
---

# Variable selection

Load the body measurements 

```{r}
load("D:/R Practice/#Classworks/#Example15/body.Rdata")
```

the research question is to explain the response which is brozek that is the percentage of the body fat calculated according to Brozek eqution.

```{r}
require(skimr)
skim_without_charts(body)
```


```{r}
pairs(body[,1:10], panel = panel.smooth)
```
```{r}
pairs(body[,7:10], panel = panel.smooth)
```

#Variance inflation factor

it required th output of the linear regression model using the full model 

```{r}
require(faraway)
modbody0 <- lm(brozek ~ ., body)
vif(modbody0)
```

we can interprete the sqrt(120.31)=10.97 saying that the standard error for the beta coefficient related to weight is around 11 times larger than it would have been with collinearity.


```{r}
sqrt(120.312)
```
we are going to calculate vif without the variable weight 

```{r}
modbody1 <-  lm(brozek ~ ., body[,-3])
vif(modbody1)
```

we set the best set of covariates using Akaike Information Criteron 

```{r}
step(modbody1)
```

```{r}
step(modbody1, direction="forward")
```

```{r}
modbody2 <- lm(brozek ~ age+height+neck+abdom+hip+thigh+wrist, data=body)
summary(modbody2)
```


